Near Optimal Behavior with Approximate State Abstraction

### 2016-02-02 paper/theory_results.tex
Reformatted all the math

It follows that the Q-value of a state $s$ in $M^T$ for action $a$ is defined by:
% zzz too long... needs to be broken up.
% zzz also, I'm not sure we want the temporally heterogeneous MDP to change state spaces mid stream. I think it should all be in terms of the ground MDP (but with the phi function used as needed)
% zzz also, ``w(g)'' comes out of the blue... there's got to be a better notation here.
\dnote{Let's put a reminder of the weighting scheme here}
\begin{equation}
Q_T(s, a) = 
\begin{cases}
	   Q_G(s, a) &  T=0\\
	   \underset{g \in X(s)}{\sum} \left[ w(g)Q_G(g,a) \right] & T = 1\\
	   \mathcal{R}_A(\epQ(s),a) + \gamma \underset{{s_A}' \in \mathcal{S}_A}{\sum} \mathcal{T}_A(\epQ(s),a,{s_A}') \max_{a'} Q_{T-1}({s_A}', a') & \text{o/w}
\end{cases}
\end{equation}

@@

% Base Case t=0
\textit{Base Case: $T = 0$}

When $T = 0$, $Q_T = Q_0$, so the result trivially follows.

% Base Case t=1
\textit{Base Case: $T = 1$}
\begin{align*}
&Q_1(s,a) = \underset{g \in X(s)}{\sum} \left[ w(g)Q_G(g,a) \right].\\
\end{align*}
Since all co-aggregated states have Q-values within $\epsilon$ of one another and $w(g)$ induces a convex combination,
\begin{align*}
&Q_1(s,a) \leq \epsilon \gamma^t + \epsilon + Q_0(s_G, a),\\
&Q_1(s,a) \leq \gamma\sum_{t=0}^{1} \epsilon \gamma^t + Q_0(s_G, a).\\
&\text{Thus,}\\
&Q_1(s,a) \left| Q_{T}(s_A, a) - Q_0(s_G,a) \right| \leq \gamma\sum_{t=0}^{T}\epsilon \gamma^t
\end{align*}

% Inductive Case
\textit{Inductive Case: $T > 1$}

@@

We denote $\epQ(s_G)$ as $s_A$.
\begin{align*}
&\text{By definition of $Q_{T}(s_A, a)$, $\mathcal{R}_A$, $\mathcal{T}_A$:}\\
&Q_T(s_A, a) =\sum_{g \in X(s_A)}w(g)\left[ R_0(g,a) + \gamma \sum_{g' \in \mathcal{S}_G} T_0(g,a,g') \max_{a'} Q_{T-1}(g', a')      \right]\\
&\text{Applying our inductive hypothesis yields}\\
&\leq \sum_{g \in X(s_A)}w(g)\left[ R_0(g,a) + \gamma \sum_{g' \in \mathcal{S}_G} T_0(g,a,g') \max_{a'}(Q_0(g', a') + \sum_{t=0}^{T-2} \epsilon \gamma^t)      \right]\\
&\leq \gamma\sum_{t=0}^{T-2} \epsilon \gamma^t + \sum_{g \in X(s_A)}w(g)\left[ R_0(g,a) + \gamma \sum_{g' \in \mathcal{S}_G} T_0(g,a,g') \max_{a'}Q_0      \right]\\
&\leq \gamma\sum_{t=0}^{T-2} \epsilon \gamma^t + \sum_{g \in X(s_A)}\left[ w(g)\ Q_0(g,a)\right]\\
&\text{Since all aggregated states have Q-values within $\epsilon$ of one another:}\\
&\leq \gamma\sum_{t=0}^{T-2} \epsilon \gamma^t + \epsilon + Q_0(s_G, a)\\
&\leq \gamma\sum_{t=0}^{T-1} \epsilon \gamma^t + Q_0(s_G, a)\\
&\text{Thus,}\\
&\left| Q_{T}(s_A, a) - Q_0(s_G,a) \right| \leq \gamma\sum_{t=0}^{T-1}\epsilon \gamma^t
\end{align*}

@@

By Equation \ref{eq:Q*Claim1}
\begin{align}
&V_G(s_G) = Q_G(s_G, a^*_G) \leq Q_A(s_G, a^*_G) + \frac{\epsilon}{1-\gamma} \\
\label{eq:Q*OptActionResult}
\end{align}

@@

Consider the policy for $M_G$ of following the optimal abstract policy $\pi^*_A$ for t steps and then following the optimal ground policy $\pi^*_G$ in $M_G$:

\begin{equation}
\pi_{A,t}(s)=
\begin{cases}
\pi_G^*(s), \text{if $t<=0$}\\
\pi_A^*(\ep(s)), \text{if $t > 0$}\\
\end{cases}
\end{equation}

For a particular $t$, the value of this policy for $s_G \in \mathcal{S}_G$ is as follows:
\begin{align*}
V^{\pi_{A,t}}(s_G) &= \\
&R_G(s, \pi_{A,t}(s_G)) + \gamma \sum_{{s_G}' \in \mathcal{S}_G}\mathcal{T}_G(s_G, a, {s_G}')V^{\pi_{A,t-1}}({s_G}')
\end{align*}

%Start induction on following the optimal abstract policy
We now show by induction on $t$ that
\begin{equation}
\forall_{t, s_G \in \mathcal{S}_g} V^{\pi^*}(s_G) \leq  V^{\pi_{A,t}}(s_G) + \sum_{i=0}^{t}\gamma^i \frac{2\epsilon}{1-\gamma}
\end{equation}

\textit{Base case: $t=0$}

@@

We assume for our inductive hypothesis that
\begin{equation}
V^{\pi^*}(s_G) \leq V^{\pi_{A,t-1}}(s_G)  + \sum_{i=0}^{t-1}\gamma^i \frac{2\epsilon}{1-\gamma}
\end{equation}

By definition 
\begin{align*}
&V^{\pi_{A,t}}(s_G) = R_G(s, \pi_{A,t}(s_G)) + \gamma \sum_{g'}\mathcal{T}_G(s_G, a, {s_G}')V^{\pi_{A,t-1}}({s_G}')\\
&\text{Applying our inductive hypothesis yields}\\
&V^{\pi_{A,t}}(g, a) \geq\\
&R_G(s, \pi_{A,t}(g, a)) + \gamma \sum_{g'}\mathcal{T}_G(g, a, g')\left(V^{\pi^*}({s_G}') - \sum_{i=0}^{t-1}\gamma^i \frac{2\epsilon}{1-\gamma} \right)\\
&\geq -\gamma\sum_{i=0}^{t-1}\gamma^i \frac{2\epsilon}{1-\gamma} + R_G(s, \pi_{A,t}(g)) + \gamma \sum_{g'}\mathcal{T}_G(g, a, g')V^{\pi^*}(g')\\
&\geq -\gamma\sum_{i=0}^{t-1}\gamma^i \frac{2\epsilon}{1-\gamma} + Q^{\pi^*}(g, \pi_{A,t} (g))\\
&\text{Applying Equation \ref{eq:Q*Claim2} yields}\\
&\geq \gamma\sum_{i=0}^{t-1}\gamma^i \frac{2\epsilon}{1-\gamma} - \frac{2\epsilon}{1-\gamma} + V_{G}(s_G)\\
&\leq \sum_{i=0}^{t}\gamma^i \frac{2\epsilon}{1-\gamma} + V_G(g)\\
& \text{Thus,}\\
&V^{\pi^*}(s_G) \leq V^{\pi_{A,t}}(s_G)  + \sum_{i=0}^{t}\gamma^i \frac{2\epsilon}{1-\gamma}
\end{align*}

Since $s_G$ was arbitrary, we conclude that our bound holds for all states in $\mathcal{S}_G$ for the inductive case.

@@

\begin{align*}
B = &\max_{s_A \in \mcS_A, s_1, s_2 \in X(s_A), a} |Q(s_1, a) - Q(s_2, a)|\\
=&\max_{s_A \in \mcS_A, s_1, s_2 \in X(s_A), a}      |R(s_1, a) - R(s_2, a) +\\
& \gamma \sum_{g' \in \mcS_G}(\mcT(s_1,a,g')-\mcT(s_2, a, g'))\max_{a'}Q(g', a')|\\
\leq &\epsilon + \gamma \sum_{s_A \in \mcS_A}\sum_{g' \in X(s_A)}\left[(T(s_1, a, g')-T(s_2, a, g')) 		\max_{a'}Q(g', a')	\right]\\
&\text{By the similarity of transitions of grouped states under $\epM$:}\\
 \leq &\epsilon + \gamma Q_{max} \sum_{s_A \in \mcS_A} \epsilon \\
\leq& \epsilon + \gamma|\mcS_G|\epsilon Q_{max}\\
&\text{Applying 0-1 reward normalization and algebra:}\\
 \leq & \frac{\epsilon + \gamma(|\mcS_G| - 1) \epsilon}{1-\gamma}
\end{align*}

>>>>

It follows that the Q-value of a state $s$ in $M^T$ for action $a$ is defined by:
% zzz too long... needs to be broken up.
% zzz also, I'm not sure we want the temporally heterogeneous MDP to change state spaces mid stream. I think it should all be in terms of the ground MDP (but with the phi function used as needed)
% zzz also, ``w(g)'' comes out of the blue... there's got to be a better notation here.
\dnote{Let's put a reminder of the weighting scheme here}
\begin{equation}
Q_T(s, a) = 
\begin{cases}
	   Q_G(s, a) &  T=0\\
	   \underset{g \in X(s)}{\sum} \left[ w(g)Q_G(g,a) \right] & T = 1\\
	   \mathcal{R}_A(\epQ(s),a) + \sigma_{T-1}(s,a) &\text{o/w}

\end{cases}
\end{equation}

Where:
\begin{equation}
\sigma_{T-1}(s,a) = \gamma \underset{{s_A}' \in \mathcal{S}_A}{\sum} \mathcal{T}_A(\epQ(s),a,{s_A}') \max_{a'} Q_{T-1}({s_A}', a')
\end{equation}

@@

% Base Case t=0
\textit{Base Case: $T = 0$}
When $T = 0$, $Q_T = Q_0$, so the result trivially follows.
% Base Case t=1
\textit{Base Case: $T = 1$}
\begin{align*}
&Q_1(s,a) = \underset{g \in X(s)}{\sum} \left[ w(g)Q_G(g,a) \right].
\end{align*}
Since all co-aggregated states have Q-values within $\epsilon$ of one another and $w(g)$ induces a convex combination,
\begin{align*}
&Q_1(s,a) \leq \epsilon \gamma^t + \epsilon + Q_0(s_G, a),\\
&Q_1(s,a) \leq \gamma\sum_{t=0}^{1} \epsilon \gamma^t + Q_0(s_G, a).
\end{align*}
Thus,
\begin{equation}
Q_1(s,a) \left| Q_{T}(s_A, a) - Q_0(s_G,a) \right| \leq \gamma\sum_{t=0}^{T}\epsilon \gamma^t
\end{equation}
% Inductive Case
\textit{Inductive Case: $T > 1$}

@@

We denote $\epQ(s_G)$ as $s_A$. By definition of $Q_{T}(s_A, a)$, $\mathcal{R}_A$, $\mathcal{T}_A$:
\begin{multline*}
Q_T(s_A, a) = \sum_{g \in X(s_A)}w(g)* \\ 
 \left[ R_0(g,a) + \gamma \sum_{g' \in \mathcal{S}_G} T_0(g,a,g') \max_{a'} Q_{T-1}(g', a')      \right]
\end{multline*}
Applying our inductive hypothesis yields,
\begin{multline*}
\leq \sum_{g \in X(s_A)}w(g) * \\ \left[ R_0(g,a) + \gamma \sum_{g' \in \mathcal{S}_G} T_0(g,a,g') \max_{a'}(Q_0(g', a') + \sum_{t=0}^{T-2} \epsilon \gamma^t)      \right]
\end{multline*}
%&\leq \gamma\sum_{t=0}^{T-2} \epsilon \gamma^t + \sum_{g \in X(s_A)}w(g)\left[ R_0(g,a) + \gamma \sum_{g' \in \mathcal{S}_G} T_0(g,a,g') \max_{a'}Q_0      \right]\\
Then:
\begin{equation*}
\leq \gamma\sum_{t=0}^{T-2} \epsilon \gamma^t + \sum_{g \in X(s_A)}\left[ w(g)\ Q_0(g,a)\right]
\end{equation*}
Since all aggregated states have Q-values within $\epsilon$ of one another:
\begin{align*}
&\leq \gamma\sum_{t=0}^{T-2} \epsilon \gamma^t + \epsilon + Q_0(s_G, a)\\
&\leq \gamma\sum_{t=0}^{T-1} \epsilon \gamma^t + Q_0(s_G, a)
\end{align*}
Thus,
\begin{equation*}
\left| Q_{T}(s_A, a) - Q_0(s_G,a) \right| \leq \gamma\sum_{t=0}^{T-1}\epsilon \gamma^t
\end{equation*}

@@

By Equation \ref{eq:Q*Claim1}
\begin{align}
&V_G(s_G) = Q_G(s_G, a^*_G) \leq Q_A(s_G, a^*_G) + \frac{\epsilon}{1-\gamma}
\label{eq:Q*OptActionResult}
\end{align}

@@

Consider the policy for $M_G$ of following the optimal abstract policy $\pi^*_A$ for t steps and then following the optimal ground policy $\pi^*_G$ in $M_G$:
\begin{equation}
\pi_{A,t}(s)=
\begin{cases}
\pi_G^*(s), \text{if $t<=0$}\\
\pi_A^*(\ep(s)), \text{if $t > 0$}
\end{cases}
\end{equation}

For a particular $t$, the value of this policy for $s_G \in \mathcal{S}_G$ is as follows:
\begin{multline*}
V^{\pi_{A,t}}(s_G) = \\
R_G(s, \pi_{A,t}(s_G)) +\ \gamma \sum_{{s_G}' \in \mathcal{S}_G}\mathcal{T}_G(s_G, a, {s_G}')V^{\pi_{A,t-1}}({s_G}')
\end{multline*}
%Start induction on following the optimal abstract policy
We now show by induction on $t$ that
\begin{equation}
\forall_{t, s_G \in \mathcal{S}_g} V^{\pi^*}(s_G) \leq  V^{\pi_{A,t}}(s_G) + \sum_{i=0}^{t}\gamma^i \frac{2\epsilon}{1-\gamma}
\end{equation}
\textit{Base case: $t=0$}

@@

We assume for our inductive hypothesis that
\begin{equation}
V^{\pi^*}(s_G) \leq V^{\pi_{A,t-1}}(s_G)  + \sum_{i=0}^{t-1}\gamma^i \frac{2\epsilon}{1-\gamma}
\end{equation}
By definition 
\begin{multline*}
V^{\pi_{A,t}}(s_G) = R_G(s, \pi_{A,t}(s_G)) + \\ \gamma \sum_{g'}\mathcal{T}_G(s_G, a, {s_G}')V^{\pi_{A,t-1}}({s_G}')
\end{multline*}
Applying our inductive hypothesis yields:
\begin{multline*}
V^{\pi_{A,t}}(g, a) \geq R_G(s, \pi_{A,t}(g, a))\ + \\ \gamma \sum_{g'}\mathcal{T}_G(g, a, g')\left(V^{\pi^*}({s_G}') - \sum_{i=0}^{t-1}\gamma^i \frac{2\epsilon}{1-\gamma} \right)
\end{multline*}
Therefore:
\begin{align*}
%&\geq -\gamma\sum_{i=0}^{t-1}\gamma^i \frac{2\epsilon}{1-\gamma} + R_G(s, \pi_{A,t}(g)) + \gamma \sum_{g'}\mathcal{T}_G(g, a, g')V^{\pi^*}(g')\\
&\geq -\gamma\sum_{i=0}^{t-1}\gamma^i \frac{2\epsilon}{1-\gamma} + Q^{\pi^*}(g, \pi_{A,t} (g))
\end{align*}
Applying Equation \ref{eq:Q*Claim2} yields:
\begin{align*}
&\geq \gamma\sum_{i=0}^{t-1}\gamma^i \frac{2\epsilon}{1-\gamma} - \frac{2\epsilon}{1-\gamma} + V_{G}(s_G)\\
&\leq \sum_{i=0}^{t}\gamma^i \frac{2\epsilon}{1-\gamma} + V_G(g)
\end{align*}

Thus,
\begin{equation*}
V^{\pi^*}(s_G) \leq V^{\pi_{A,t}}(s_G)  + \sum_{i=0}^{t}\gamma^i \frac{2\epsilon}{1-\gamma}
\end{equation*}
Since $s_G$ was arbitrary, we conclude that our bound holds for all states in $\mathcal{S}_G$ for the inductive case.

@@

\begin{align*}
B = &\max_{s_A \in \mcS_A, s_1, s_2 \in X(s_A), a} |Q(s_1, a) - Q(s_2, a)|\\
=&\max_{s_A \in \mcS_A, s_1, s_2 \in X(s_A), a}      |R(s_1, a) - R(s_2, a) +\\
& \gamma \sum_{g' \in \mcS_G}(\mcT(s_1,a,g')-\mcT(s_2, a, g'))\max_{a'}Q(g', a')|
\end{align*}
\begin{multline*}
\leq \epsilon + \gamma \sum_{s_A \in \mcS_A}\sum_{g' \in X(s_A)} \\ \left[(T(s_1, a, g')-T(s_2, a, g')) 		\max_{a'}Q(g', a')	\right]
\end{multline*}
By the similarity of transitions of grouped states under $\epM$:
\begin{align*}
 \leq &\epsilon + \gamma Q_{max} \sum_{s_A \in \mcS_A} \epsilon \\
\leq& \epsilon + \gamma|\mcS_G|\epsilon Q_{max}\\
&\text{Applying 0-1 reward normalization and algebra:}\\
 \leq & \frac{\epsilon + \gamma(|\mcS_G| - 1) \epsilon}{1-\gamma}
\end{align*}
